Arc Blanc: a real time ocean simulation framework

### 2024-11-14 paper.tex
Fix plural
Add capital letter for footnote...

%-------------------------------------------------------------------------
\section{Ocean waves}
\label{sec:oceanWaves}
This section summarizes Tessendorf's method and the various improvements it has received.
It also presents our first contributions: calculation of the fluid velocity at any depth.

Let us introduce some notations as follows:
\begin{itemize}
	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water. \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{see appendix \ref{sec:annexe} for more details on this assumption.}.
	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x ~tz\right]^T$ and at time $t$.
	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $\phi(\mathbf x,y,t)\in\mathbb R$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
\end{itemize}

Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{see appendix \ref{sec:annexe-NS}.}:
\begin{equation}
	\label{equa:systemTessendorf}
	\left\lbrace
	\begin{array}{ll}
		\frac{\partial\phi}{\partial t} = -gh(\mathbf x,t)            & \text{for }y=0,            \\
		\Delta\phi=0                                                  & \text{for }-H\leq y\leq 0, \\
		\frac{\partial h}{\partial t}=\frac{\partial\phi}{\partial y} & \text{for }y=0,            \\
		\frac{\partial\phi}{\partial y}=0                             & \text{for }y=-H.
	\end{array}
	\right.
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{see appendix \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Incompressibility}
As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal\footnote{see \url{https://en.wikipedia.org/wiki/Sound_speed_profile} to $c_{sound} = 1530~m.s^{-1}$.}.
Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.
More formally:
\begin{equation}
	\rho(t)=\rho.
\end{equation}
Consequently, density can be factorized in each term of equations \ref{equa:ns0}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Irrotational}
If a fluid is irrotational at time $t_0$, it must remain irrotational unless there are changes in the boundary conditions.
Clearly, flat sea are irrotational fluid, corresponding to $\nabla\times\mathbf{u} = \mathbf{0}$.
Therefore, it can assume that for calm waves the fluid remains irrotational.
Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{see \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.

>>>>

%-------------------------------------------------------------------------
\section{Ocean waves}
\label{sec:oceanWaves}
This section summarizes Tessendorf's method and the various improvements it has received.
It also presents our first contribution: calculation of the fluid velocity at any depth.

Let us introduce some notations as follows:
\begin{itemize}
	\item The scalar $H\in\mathbb R^+$ denotes the depth of the water. \Framework assumes \textbf{deep water} and therefore supposes that $H$ tends towards $+\infty$\footnote{See appendix \ref{sec:annexe} for more details on this assumption.}.
	\item The scalar $h(\mathbf x, t)\in\mathbb R$ is the height of the water surface at the horizontal position $\mathbf x=\left[x ~tz\right]^T$ and at time $t$.
	\item The vector $\mathbf{v}(\mathbf x, y, t) =  \left[v_x(\mathbf x, y, t)\ v_y(\mathbf x, y, t)\ v_z(\mathbf x, y, t)\right]^T\in\mathbb R^3$ represents the velocity of the fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $\phi(\mathbf x,y,t)\in\mathbb R$ is the potential velocity of fluid at the horizontal position $\mathbf x$, depth $y$ and time $t$.
	\item The scalar $g=9.80665~m.s^{-2}$ denotes the standard gravity, while $\mathbf g\in\mathbb R^3$ is the gravitational field.
\end{itemize}

Tessendorf method is based on the following approximations of the Navier-Stokes equations\footnote{See appendix \ref{sec:annexe-NS}.}:
\begin{equation}
	\label{equa:systemTessendorf}
	\left\lbrace
	\begin{array}{ll}
		\frac{\partial\phi}{\partial t} = -gh(\mathbf x,t)            & \text{for }y=0,            \\
		\Delta\phi=0                                                  & \text{for }-H\leq y\leq 0, \\
		\frac{\partial h}{\partial t}=\frac{\partial\phi}{\partial y} & \text{for }y=0,            \\
		\frac{\partial\phi}{\partial y}=0                             & \text{for }y=-H.
	\end{array}
	\right.
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Velocity of water}
\label{subsec:velWater}
Starting from the system \ref{equa:systemTessendorf}, \framework proposes to compute the velocity of the water at any depth $y$, as follows\footnote{See appendix \ref{sec:annexe} for more details on the origin of this expression}:
\begin{equation}\label{equa:velocity}
	% \left[
	%   \begin{array}{c} \displaystyle v_x(\mathbf x,y,t) \\ \displaystyle v_y(x,y,z,t) \\ \displaystyle v_z(x,y,z,t) \end{array}
	% \right]
	\mathbf v(\mathbf x,y,t)=
	\sum_{\mathbf{k}} \tilde{\mathbf{v}}(\mathbf{k},t,y)\exp\left(i\mathbf{k}\cdot\mathbf x\right),
\end{equation}
where the vector $\tilde{\mathbf{v}}(\mathbf{k},t, y)$ is given as follows:
\begin{equation}
	\tilde{\mathbf{v}}\left(\mathbf{k},t,y\right)=%
	E\left(\mathbf{k},y\right)\left(\tilde{h}_0\left(\mathbf{k}\right)\exp\left({i\omega(k)t}\right)-\tilde{h}_0^*\left(-\mathbf{k}\right)\exp\left({-i\omega(k)t}\right)\right)
	\left[
		\begin{array}{c}
			\displaystyle\frac{-k_xg}{\omega(k)} \\
			\displaystyle i\omega(k)             \\
			\displaystyle\frac{-k_zg}{\omega(k)}
		\end{array}
		\right],
\end{equation}
where the attenuation function $E(\mathbf{k},y)\in\mathbb R$ is defined as follows:
\begin{equation}
	\label{equa:attenuationDeep}
	E(\mathbf{k}, y)=
	\begin{cases}
		1+ky     & \text{if }y>0, \\
		\exp(ky) & \text{else}.
	\end{cases}
\end{equation}
The first case ($y>0$) comes from an extrapolation above the sea surface of the normal case ($y\leq0$).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Incompressibility}
As mention in Section~3.3.2 of Hulin \etal (2001) \cite{hulinHydrodynamiquePhysique2001}, a fluid can be considered incompressible if $u<<c_{sound}$ where $c_{sound}$ if the sound speed in ocean which is approximately equal\footnote{See \url{https://en.wikipedia.org/wiki/Sound_speed_profile} to $c_{sound} = 1530~m.s^{-1}$.}.
Aside from phenomena such as hurricane, this inequality seems reasonable for oceans, and leads to consider that ocean is incompressible.
More formally:
\begin{equation}
	\rho(t)=\rho.
\end{equation}
Consequently, density can be factorized in each term of equations \ref{equa:ns0}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Irrotational}
If a fluid is irrotational at time $t_0$, it must remain irrotational unless there are changes in the boundary conditions.
Clearly, flat sea are irrotational fluid, corresponding to $\nabla\times\mathbf{u} = \mathbf{0}$.
Therefore, it can assume that for calm waves the fluid remains irrotational.
Then, there exists a scalar field called \textbf{velocity potential} defined by $\phi = \mathbf{\nabla} \phi$ from \emph{Helmholtz decomposition}\footnote{See \url{https://en.wikipedia.org/wiki/Helmholtz_decomposition}.}.
